---
output: pdf_document
---
# An accessment : How do you know you exercise correctly ? 

Background
========

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

Dataset
========

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

Download both training dataset :-
```{r}

set.seed(11111)
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)

curdir <-getwd()
file.url<-'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
download.file(file.url,destfile=paste(curdir,'/pml-training.csv',sep=""))

curdir <-getwd()
file.url<-'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(file.url,destfile=paste(curdir,'/pml-testing.csv',sep=""))

```

Load both dataset :-

```{r}
train <- read.csv(paste(curdir,'/pml-training.csv',sep=""),na.strings=c("NA","#DIV/0!",""))
test <-read.csv(paste(curdir,'/pml-testing.csv',sep=""),na.strings=c("NA","#DIV/0!",""))
```


Cheking the dimension of training and test dataset :-

```{r}
dim(train)
dim(test)
```

Checking the columns which have all missing values

```{r}
train<-train[,colSums(is.na(train)) == 0]
test <-test[,colSums(is.na(test)) == 0]
```

We remove 6 of the variables which is irrelevant like :-

a)  user_name

b)  raw_timestamp_part_1

c)  raw_timestamp_part_2

d)  cvtd_timestamp

e)  new_window

f)  num_window 


which resides on the column 1-7.

```{r}
train <-train[,-c(1:7)]
test <-test[,-c(1:7)]
```

Check again the dimension

```{r}
dim(train)
dim(test)
```

Now we obtain the several rows to preview

```{r}
head(train)
head(test)
```

In order to run cross-validation , the training dataset need to partition into 2 sets . We set the 1st partition for training dataset to 75% and test dataset to 25%. Training dataset contains 53 variables with 19622 obs and test dataset contains 53 variables with 20 obs.

This will do the randomize sub-sampling without replacement 


```{r}
PartTrain <- createDataPartition(y=train$classe, p=0.75, list=FALSE)
train_part <- train[PartTrain, ]; 
test_part <- train[-PartTrain, ]
dim(train_part)
dim(test_part)

```

Visualization
============
We try to plot into the histogram to see the trending frequency of each sub-training & test dataset by comparing with each other. The variable classe contains 5 levels which is A,B,C,D & E


```{r}
plot(train_part$classe, col="lightgreen", 
main="( train_part ) - Classe vs. Frequency", 
xlab="Classe", ylab="Frequency")
```

The graph above shows that A ~ 4000x occurrences is most frequent while D is the lest frequent ~ 2500x occurrences


Decision Tree
============
We try to plot



Reference
============
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3lj0hACeI





